### 2018-12-22 introduction.tex
finished arot considerations, rwy usage.
began LTI method

Keflavík International Airport (IATA:KEF, ICAO:BIKF) is the major gateway to Iceland currently with two functioning runways. The runways operate in both directions and are designated as RWY-01 (South-North), RWY-19 (North-South) , RWY-10 (East-West) and RWY-28 (West-East). 
There has been a gradual increase of passengers travelling through Keflavík Airport in the recent years with over eight million passengers and 62,3 thousand air transport movements in 2017~\cite{isavia_facts_2017}. For 2018 the passenger numbers show a growth of 13.1\% on average (at the time of writing) from the previous year~\cite{isavia_pass_statistics_2018}. The increase in traffic at Keflavík Airport can encounter constrained capacity of the runways during peak traffic periods. 
A significant constraint of airport capacity is caused by separation regulations of aircraft due to wake-vortices. These are specified by rules established by the European Aviation Safety Agency (EASA) for each type of aircraft based primarily on the weight of the aircraft. The general flight safety requirements are established by the International Civil Aviation Organisation (ICAO) for maintaining safe distance between aircraft.\fxnote{Introduction needs more work, incomplete.}

@@

\fxnote{describe rot in general and what factors have influence on rot, theory, models for rot, see validation for rot model paper}

>>>>

Keflavík International Airport (IATA:KEF, ICAO:BIKF) is the major gateway to Iceland currently with two functioning runways. The runways operate in both directions and are designated as RWY-01 (South-North), RWY-19 (North-South) , RWY-10 (West-East) and RWY-28 (East-West). 
There has been a gradual increase of passengers travelling through Keflavík Airport in the recent years with over eight million passengers and 62,3 thousand air transport movements in 2017~\cite{isavia_facts_2017}. For 2018 the passenger numbers show a growth of 13.1\% on average (at the time of writing) from the previous year~\cite{isavia_pass_statistics_2018}. The increase in traffic at Keflavík Airport can encounter constrained capacity of the runways during peak traffic periods. 
A significant constraint of airport capacity is caused by separation regulations of aircraft due to wake-vortices. These are specified by rules established by the European Aviation Safety Agency (EASA) for each type of aircraft based primarily on the weight of the aircraft. The general flight safety requirements are established by the International Civil Aviation Organisation (ICAO) for maintaining safe distance between aircraft.\fxnote{Introduction needs more work, incomplete.}

@@
